---
title: "Introduction to tidymodels"
author: Max Kuhn
event: New York R Conference
url: https://github.com/topepo/2022-nyr-workshop
output:
  xaringan::moon_reader:
    anchor_sections: FALSE
    css: ["default", "css/theme.css", "css/fonts.css"]
    seal: false 
    lib_dir: libs
    nature:
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
      highlightStyle: rainbow
      highlightLanguage: ["r", "css", "yaml"]
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
---

```{r include = FALSE}
# devtools::install_github("gadenbuie/countdown")
library(countdown)
library(leaflet)
library(knitr)
library(ongoal)

knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  digits = 3, 
  fig.path = "images/introduction-",
  fig.align = 'center',
  fig.width = 10,
  fig.height = 6,
  out.width = "95%",
  dev = 'svg',
  dev.args = list(bg = "transparent"),
  tidy = FALSE
)

opts_chunk$set()
source("_common.R")
```

class: title-slide, left, middle
background-image: url("images/tidymodels.svg")
background-position: 85% 50%
background-size: 30%
background-color: #F9F8F3

.pull-left[

# `r rmarkdown::metadata$title`

## `r rmarkdown::metadata$author`

### `r rmarkdown::metadata$event`

### Repo: `r slide_url`

]


---
# Our goals for this workshop

* Introduce tidymodels and its general philosophy on modeling. 

* Help you become proficient with the core packages for modeling. 

* Point you to places to learn more and get help. 

---
# Why tidymodels?

There are several other modeling frameworks in R that try to:

 * create a uniform, cohesive, and unsurprising set of modeling APIs
 
Examples are `r pkg(caret)`, `r pkg(mlr3)`, and others. 

* `r pkg(caret)` is more favorable for people who prefer base R/traditional interfaces.

* tidymodels would probably be preferable to those who place importance on a tidy _R_ interface, a large number of features, and the idea that the interfaces should enable the "pit of success". 

* `r pkg(mlr3)` is more pythonic and also has many features.

---
# The tidymodels package

There are a lot of tidymodels packages but about 90% of the work is done by 5 packages. 

The best way to get started with tidymodels is to use the `r pkg(tidymodels)` meta-package. It loads the core packages plus some tidyverse packages.

Some helpful links:

 * List of [all tidymodels functions](https://www.tidymodels.org/find/#search-all-of-tidymodels)
 * List of [all parsnip models](https://www.tidymodels.org/find/parsnip/)
 * List of [all recipe steps](https://www.tidymodels.org/find/recipes/)


---
# The tidymodels package `r I(hexes(c("tidymodels", "tidyverse")))`

```{r}
library(tidymodels)
```

---
# Managing name conflicts

```r
tidymodels_prefer(quiet = FALSE)
```

```
## [conflicted] Will prefer dplyr::filter over any other package
## [conflicted] Will prefer dplyr::select over any other package
## [conflicted] Will prefer dplyr::slice over any other package
## [conflicted] Will prefer dplyr::rename over any other package
## [conflicted] Will prefer dials::neighbors over any other package
## [conflicted] Will prefer plsmod::pls over any other package
## [conflicted] Will prefer purrr::map over any other package
## [conflicted] Will prefer recipes::step over any other package
## [conflicted] Will prefer themis::step_downsample over any other package
## [conflicted] Will prefer themis::step_upsample over any other package
## [conflicted] Will prefer tune::tune over any other package
## [conflicted] Will prefer yardstick::precision over any other package
## [conflicted] Will prefer yardstick::recall over any other package
```


---
# Base R and tidyverse differences

.pull-left[
Base R/`r pkg(caret)`
```{r, eval = FALSE}
mtcars <- mtcars[order(mtcars$cyl),]
mtcars <- mtcars[, "mpg", drop = FALSE]


# ──────────────────────────────────────────────

mtcars$mp        # matches incomplete arg
mtcars[, "mpg"]  # a vector


# ──────────────────────────────────────────────

num_args <- function(x) length(formals(x))


num_args(caret::trainControl) + 
  num_args(caret:::train.default)

```

```
## 38
```
]
.pull-right[
tidyverse/tidymodels
```{r, eval = FALSE}
mtcars %>% 
  arrange(cyl) %>% 
  select(mpg)

# ──────────────────────────────────────────────

tb_cars <- as_tibble(mtcars)
tb_cars$mp        # fails
tb_cars[, "mpg"]  # A tibble 

# ──────────────────────────────────────────────



num_args(linear_reg) + num_args(set_engine) + 
  num_args(tune_grid) + num_args(control_grid) + 
  num_args(vfold_cv)

```
```
## 23
```
]


---
# Example data set

We'll use some sports data to demonstrate classification models. 

The data are in the ongoal package at [`topepo/ongoal`](https://github.com/topepo/ongoal).

The outcome is a factor of whether a shot was on-goal or not. Predictors:

* Game related: type of game, power-play data, etc.
* Player: who, position, team
* Shot: where, when

We'll use data from the 2015-2016 season. 

---

# Data ~~Slapshot~~ Snapshot

```{r rink, fig.width=7, fig.height=3, out.width = "80%"}
library(ongoal)
# source("_common.R")
set.seed(1)
on_goal %>%
  filter(season == "20152016") %>% 
  sample_n(500) %>% 
  plot_rink(alpha = 3/4, cex = 2)
```

.footnote[ggplot rink code by `mrbilltran`]

---

# Medium distance shots by player position

```{r position, fig.width=7, fig.height=3, out.width = "80%"}
# Change the grouping variable
set.seed(1)
on_goal %>%
  filter(season == "20152016" & distance > 32) %>% 
  sample_n(500) %>% 
  plot_rink(position, alpha = 3/4, cex = 2)
```


---

# Hands-On: Explore the Data

Take a look at these data for a few minutes and see if you can find any interesting characteristics in the predictors or the outcome.  

`r countdown(minutes = 10, left = 1)`


---
layout: false
class: inverse, middle, center

# [`tidymodels.org`](https://www.tidymodels.org/)

# _Tidy Modeling with R_ ([`tmwr.org`](https://www.tmwr.org/))




